Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain

### 2018-02-19 pkg/disttree/inst/draft/draft.Rnw
changed format of references and revised parts of the application and methodology section
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2030 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: fe018959bfd40c6c95a1abd42210653c4f28fc5d

\Abstract{
While many regression models consider only the mean of the response variable recent advances in parametric modeling have made it possible to model all distribution parameters. In that way the underlying distribution can be fully specified allowing for a wide range of analysis and prediction methods which are particularly valuable in applications such as probabilistic weather forecasting. In generalized additive models for location, scale, and shape %(GAMLSS, \citealt{stasinopoulos2005}) 
each distribution parameter can be modeled separately by its own generalized additive model (GAM) capturing linear and non-linear effects. A different approach that would make it possible to also model non-additive effects and interactions is to apply tree-structured models such as regression trees %(\citealt{loh2011classification}) 
or random forests. %(\citealt{breiman2001random}). 
%%%<TH> "random forest" generisch; Breiman and Cutler's randomForest </TH>
This class of tree and forest models will now be integrated in probabilistic forecasting providing a novel general distributional framework for these methods. In particular, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. 
%%%<TH> break-up very long sentence </TH>
Applying these novel approaches to precipitation data in a mountainous region with a very large number of numerical weather prediction quantities shows one of their main advantages in selecting variables automatically and illustrates their strengths as their performance is compared to GAMLSS models that are specified based either on subject-matter knowledge and experience from meteorology or based on a computationally more demanding boosting approach.
}
\Keywords{parametric models, regression trees, random forests, recursive partitioning, probabilistic forecasting, distributional regression}

@@

In regression analysis a wide range of models has been developed to describe the
relationship between a response variable and a set of covariates. The classical
model is the linear model (LM) where the conditional mean of the response
is modeled through a linear function of the covariates (see the left panel of
Figure~\ref{Devel_parmod} for a schematic illustration). Over the last decades
this has been extended in various directions including:
\begin{itemize}
  \item \emph{Generalized linear models} (GLMs, \citealp{Nelder+Wedderburn:1972})
    encompassing an additional nonlinear link function for the conditional mean.
  \item \emph{Generalized additive models} (GAMs, \citealp{Hastie+Tibshirani:1986})
    allowing for smooth nonlinear effects in the covariates
    (Figure~\ref{Devel_parmod}, middle).
  \item \emph{Generalized additive models for location, scale, and shape}
    (GAMLSS, \citealt{stasinopoulos2005}) adopting a probabilistic modeling
    approach. In GAMLSS, each parameter of a statistical distribution can depend
    on an additive predictor of the covariates comprising linear and/or
    smooth nonlinear terms (Figure~\ref{Devel_parmod}, right).
\end{itemize}
Thus, the above-mentioned models provide a broad toolbox for capturing different
aspects of the response (mean only vs.\ full distribution) and different types
of dependencies on the covariates (linear vs.\ nonlinear additive terms).

@@

While in many applications conditional mean regression models have been receiving
the most attention, there has been a paradigm shift over the last decade towards
distributional regression models. An important reason for this is that in many
fields forecasts of the mean are not the only (or not even the main) concern but
instead there is an increasing interest in probabilistic forecasts. Quantities of
interest typically include exceedence probabilities for certain thresholds of the
response or quantiles of the response distribution. As an example, consider
weather forecasting where there is less interest in the mean amount of
precipitation on the next day. Instead, the probability of rain vs.\ no rain
is typically more relevant or, in some situations, a prediction interval of
expected precipitation (say from the expected 10\% to 90\% quantile). Similar
considerations apply for other meteorological quantities and hence attention
in the weather forecasting literature has been shifting from classical linear
models \citep{Glahn+Lowry:1972} towards probabilistic models such as the
non-homogeneous Gaussian regression (NGR) of \cite{gneiting2005calibrated}.
The NGR typically describes the mean of some meteorological response variable through
the average of the corresponding quantity from an ensemble of physically-based
numerical weather predictions (NWPs). Similarly, the variance of the response
is captured through the variance of the ensemble of NWP outputs. Thus, the NGR
considers both the mean as well as the uncertainty of the NWP model outputs
to obtain probabilistic forecasts calibrated to a particular site.

%% late: introduce MOS and EMOS? more details about precipitation forecasting
%In ensemble model output statistics (EMOS, \cite{gneiting2005calibrated}) this idea is taken one step further. Usually acting on the assumption of a normally distributed response variable this model allows for both parameters, the mean and the variance, to depend linearly on covariates provided by NWPs. However, the restriction to model only linear effects has risen the interest in applying more flexible models such as offered by the GAMLSS framework.

@@

In summary, the models discussed so far provide a broad and powerful toolset for
parametric distributional fits depending on a specified set of additive linear or
smooth nonlinear terms. A rather different approach to capturing the dependence on
covariates are tree-based models.
\begin{itemize}
  \item \emph{Regression trees} (\citealt{breiman1984classification}) recursively
    split the data into more homogeneous subgroups and can thus capture abrupt shifts
    (Figure~\ref{Devel_treeforest}, top left) and approximate nonlinear functions.
    Furthermore, trees automatically carry out a forward selection of covariates and
    their interactions. 
  \item \emph{Random forests} (\citealt{breiman2001random}) average the predictions of
    an ensemble of trees fitted to resampled versions of the learning data. This
    stabilizes the recursive partitions from individual trees and hence better
    approximates smooth functions (Figure~\ref{Devel_treeforest}, top middle)
\end{itemize}
While classical regression trees and random forests only model the mean of the response
we propose to follow the ideas from GAMLSS modeling -- as outlined
in Figure~\ref{Devel_parmod} -- and combine tree-based methods with parametric distributional models,
yielding two novel techniques:
\begin{itemize}
  \item \emph{Distributional trees} split the data into more homogeneous groups with
    respect to a parametric distribution, thus capturing changes in any distribution
    parameter like location, scale, or shape (Figure~\ref{Devel_treeforest}, bottom left)
  \item \emph{Distributional forests} utilize an ensemble of distributional trees
    for obtaining stabilized and smoothed parametric predictions (Figure~\ref{Devel_treeforest},
    top right).
\end{itemize}
In the following, particular focus is given to distributional forests as a method for obtaining
probabilistic forecasts by leveraging the strengths of random forests: the ability to
capture both smooth and abruptly changing functions along with simultaneous selection of
variables and possibly complex interactions. Thus, these properties make the method particularly
appealing in case of many covariates with unknown effects where it would be challenging
to specify a distributional regression model like GAMLSS.

In weather forecasting, these properties are appealing in mountainous regions
and complex terrain where small-scale effects are not well-resolved in the NWP
outputs. As these are computed at a coarser scale, there may be abrupt local
effects within a single NWP grid cell. To illustrate this in practice,
precipitation forecasts are obtained with distributional forests at 95
meteorological stations in a mountainous region in the Alps, covering mainly
Tyrol, Austria, and adjacent areas. More specifically, a zero-censored Gaussian
distribution is employed to model 24-hour total precipitation so that the
zero-censored point mass describes the probability of no precipitation on a
given day (see Figure~\ref{axams_onestation}). Forecasts for July are
established based on data from the same month in the years 1985--2012 along with
82~covariates derived from a wide range of NWP ensemble outputs. As
Figure~\ref{axams_onestation} shows, the station-wise models yield a full
distributional forecast for each day -- here for the same day (July~24) at one
station (Axams) over four years (2009--2012) -- based on the previous 24~years
as training data. The corresponding observations conform reasonably well with
the predictions. In Section~\ref{ProbabilisticForecasting} we investigate the
performance of distributional forests in this forecasting task in more detail.
It is shown that they perform at least on par and sometimes clearly better than
three alternative zero-censored Gaussian models: a standard ensemble model
output statistics approach (EMOS, \citealp{gneiting2005calibrated}), a GAMLSS
with regressors prespecified based on meteorological expertise following
\cite{Stauffer+Umlauf+Messner:2017}, and a boosted GAMLSS
\citep{gamboostLSS2016} using non-homogeneous boosting
\citep{Messner+Mayr+Zeileis:2017} as a technique for variable selection from all
82~available regressors.

@@

\subsection{Motivation}
As illustrated in Figure~\ref{Devel_parmod} a lot of development has been made in the field of parametric modeling and a wide range of tools can be applied in research areas such as probabilistic forecasting. But as mentioned before, further features such as the ability to capture non-additive effects or interactions are desirable in many applications which demands for more flexible models. Since tree-structured models provide these required features a combination of them with parametric models can offer a framework that is able to deal with a wide range of different modeling situations. In this section we are going to introduce the novel methodology which embeds parametric modeling into the idea of tree-structured models step by step. We start with a simple distributional fit in Section~\ref{distfit}, then build a distributional tree in Section~\ref{disttree} and finally a distributional forest in Section~\ref{distforest}.

@@

The goal is to fit a distributional model $D(Y, \theta)$ to a response variable $Y \in \mathcal{Y}$ for a \textit{k}-dimensional vector of distribution parameters $\theta \in \Theta$ in case of a \textit{k}-parametric distribution. The distribution family has to be specified in advance such that a log-likelihood function $\ell(\theta; Y)$ is provided. Using the GAMLSS framework (\citealt{stasinopoulos2005}) a wide range of distributions can be selected where each distribution parameter can be specified separately. This enables the model to deal with various features that can usually be very challenging in model fitting such as censoring, truncation or heavy tails.

In our application of precipitation forecasting the response variable, which is the power transformed total precipitation amount within 24 hours, is expected to be normally distributed with mean/location parameter $\mu$ and standard deviation/scale parameter $\sigma$ and left-censored with censoring point 0. (The power transformation is used to remove large parts of the skewness of the data as further explained in Section~\ref{data}.) Therefore, the corresponding log-likelihood function with parameter vector $\theta = (\mu, \sigma)$ is
$$
\ell(\mu, \sigma; Y) = 
\begin{cases}
    \log\left(\frac{1}{\sigma} \cdot \phi\left(\frac{Y - \mu}{\sigma}\right) \right), & \text{if } Y > 0\\
    \vspace{0.1cm}
    \log\left(\Phi\left(\frac{-\mu}{\sigma}\right)\right), & \text{if } Y = 0
\end{cases}
$$
where $\phi$ and $\Phi$ are the probability density function and the distribution function of the standard normal distribution $\mathcal{N}(0,1)$. Next to the censored normal distribution the censored logistic distribution is also often chosen for the modeling of precipitation amounts. In \citealt{Gebetsberger2017} this choice has been made in order to be able to deal with heavy tails which often appear in precipitation data. Another frequent application of probabilistic forecasting in meteorology is the modeling of near surface air temperature. In this case a normal distribution is often selected.

@@

A common way of including these covariates is to consider them as regressors in a distributional model such that each parameter $\theta$ depends on covariates $\bold{Z} = (Z_1,\ldots,Z_m)$. This could for example be done by a linear function $g$. In that way the fitted model would be specified as $D(Y, \theta = g(\bold{Z}))$. In order to avoid the restriction to linear effects one might also use a set of smooth functions $\{f_i\}_{i = 1,\ldots,p}$ to specify a more flexible model $D(Y, \theta = f_1(\bold{Z}) + \ldots + f_p(\bold{Z}))$ applying a GAM structure. Another approach to distributional regression is proposed by \citealt{Klein2015} where each parameter of a possibly complex distribution can be modeled by an additive term consisting of a variety of different functional effect types. These include non-linear effects, spatial effects, random coefficients and interaction surfaces which extends the flexibility compared to parametric models such as GAMLSS modeling only smooth effects.

@@

\subsection{Distributional forest}
\label{distforest}
A distributional forest consists of an ensemble of $T$ distributional trees. Each of these trees is built on a different data set which can either be a bootstrap or a subset of the original data set. Moreover, in each node a subset of the partitioning variables is chosen randomly as candidates for splitting variables. In that way the correlation among the resulting trees is reduced and therefore also the variance of the model (\citealt{breiman2001random}).

Forest models differ mainly in the way how the trees are combined to make predictions. Following the strategy explained and proposed by \cite{transformationforests2017} we interpret random forests as adaptive local likelihood estimators as distributional models are fit locally in subgroups while the term ``adaptive'' refers to the dependence on the splitting variables $\bold{Z}$ and the resulting subgroups and weighing schemes. In this case the forest itself is only used to evaluate neighborhood relationships among the observations which can then be included in the fitting process. In transformation forests by \cite{transformationforests2017} so-called ``nearest neighbor weights'' are calculated as a measure of distance between observations. This approach has also already been applied in several random forest-type methods such as in ``adaptive nearest neighbors'' by \cite{lin2006}, in ``quantile regression forests''  by  \cite{meinshausen2006} and in ``bagging survival trees'' by \cite{hothorn2004}. The idea of nearest neighbor weights is to measure similarity of a pair of observations by counting the number of trees in which it is asigned to the same terminal node.
%Similar observations have a high probability of ending up in the same terminal node whereas this probability is low for quite different observations. 
%although the first random forest-type algorithm for the estimation of conditional distribution functions was published more than a decade ago (bagging survival trees, Hothorn et al. 2004). 
In particular, for a new set of covariates $\bold{z} = (z_1, \ldots, z_m) $ such a set of weights 
$\bold{w}^{\text{forest}}~=~(w^{\text{forest}}_1, \ldots, w^{\text{forest}}_n)$ is obtained by the following procedure: For each tuple $(y_i,\bold{z}_i)$ in the learning data set $\{(y_i,\bold{z}_i)\}_{i=1\ldots,n}$ a weight is calculated by counting the number of trees in which $(y_i,\bold{z}_i)$ ends up in the same terminal node as the new set of covariates $\bold{z}$:
$$
w^{\text{forest}}_i(\bold{z}) = \sum_{t=1}^T \sum_{b=1}^{B^t} I((\bold{z}_i \in \mathcal{B}^t_b) \land (\bold{z} \in \mathcal{B}^t_b))
$$
where $T$ is the number of trees of the forest, $B^t$ is the number of terminal nodes of the \textit{t}-th tree and $\mathcal{B}^t_b$ represents the \textit{b}-th terminal node of the \textit{t}-th tree.

@@

As explained in the introduction, distributional forests provide a new approach in probabilistic forecasting by embedding the ideas of flexible parametric modeling into the framework of tree-structured models. The ability to deal with a wide variety of effects and an automatic variable selection and detection of interactions are features provided by distributional forests that could be particularly valuable in probabilistic (weather) forecasting in order to tackle the above mentioned challenges.
To show their strengths in practice, distributional forests are now applied to probabilistic precipitation forecasts in a mountainous region based on a large number of numerical weather prediction quantities. Their results are compared to those of other zero-censored Gaussian models including a standard ensemble model output approach and two GAMLSS models. For one of the GAMLSS models the regressors have been selected and interactions have been specified in advance based on subject-matter knowledge and experience from meteorology. The other one is based on a computationally more demanding boosting approach.

@@

Observations are available on a daily basis, each representing the total precipitation sum fallen within 24~hours observed at 6~UTC at 95 stations all over Tyrol and its close surrounding. The observations are undergoing a rigorous quality check before made available to the public.  As covariates the second generation reforecast data set of the global ensemble forecast system (GEFS) is used (\citealt{hamill2013}).  This data set consists of an 11-member ensemble based on a fixed version of the numerical model and a horizontal grid-spacing of about $50~\times~50 \text{km}^2$ initialized daily at 0~UTC from December 1984 to present providing forecasts on 26 hourly temporal resolution. Each of the 11 ensemble members uses slightly different perturbed initial conditions trying to predict the situation specific uncertainty of the atmospheric state.

@@

%%% <TH> hm, now we _are_ using a transformation model without estimating the
%%% transformation. </TH>
To remove large parts of the skewness of precipitation data, a power transformation has been used frequently (Box and Cox 1964). In literature, cubic (\citealt{stidd1973}) or square root (\citealt{hutchinson1998b}) transformations have often been suggested but may vary for different climatic zones or temporal aggregation periods. In \citealt{stauffer2017a} optimization methods have been applied to find the best transformation factor which resulted in a value of $\frac{1}{1.6}$. Since the observation data set used for this optimization contains the observations considered in this article the same factor is chosen here for the power transformation of the response variable \emph{tp} (total precipitation). The resulting transformed variable is expected to be normally distributed and left-censored at censoring point 0. For simplicity the same transformation has been chosen for the variable \emph{cape} (convective available potential energy).

@@

\item An ensemble model output statistics (EMOS, \citealp{gneiting2005calibrated}) model where the location parameter and the logarithm of the scale parameter depend linearly on the ensemble mean forecast of the total daily precipitation sum and the corresponding ensemble standard deviation respectively.
The choice of the logarithmic function as link function for the scale parameter has been made based on the results of \citealt{Gebetsberger2017} where different link functions have been compared. Moreover, testing the identity, quadratic and logarithmic function as link functions in this setting supported this choice as the logarithmic function led to the best performance of the EMOS model in this application.

>>>>

\Abstract{
While many regression models consider only the mean of the response variable recent advances in parametric modeling have made it possible to model all distribution parameters. In that way the underlying distribution can be fully specified allowing for a wide range of analysis and prediction methods which are particularly valuable in applications such as probabilistic weather forecasting. In generalized additive models for location, scale, and shape %(GAMLSS, \citealt{Rigby+Stasinopoulos:2005}) 
each distribution parameter can be modeled separately by its own generalized additive model (GAM) capturing linear and non-linear effects. A different approach that would make it possible to also model non-additive effects and interactions is to apply tree-structured models such as regression trees %(\citealt{Loh:2011}) 
or random forests. %(\citealt{Breiman:2001}). 
%%%<TH> "random forest" generisch; Breiman and Cutler's randomForest </TH>
This class of tree and forest models will now be integrated in probabilistic forecasting providing a novel general distributional framework for these methods. In particular, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. 
%%%<TH> break-up very long sentence </TH>
Applying these novel approaches to precipitation data in a mountainous region with a very large number of numerical weather prediction quantities shows one of their main advantages in selecting variables automatically and illustrates their strengths as their performance is compared to GAMLSS models that are specified based either on subject-matter knowledge and experience from meteorology or based on a computationally more demanding boosting approach.
}
\Keywords{parametric models, regression trees, random forests, recursive partitioning, probabilistic forecasting, distributional regression}

@@

In regression analysis a wide range of models has been developed to describe the
relationship between a response variable and a set of covariates. The classical
model is the linear model (LM) where the conditional mean of the response
is modeled through a linear function of the covariates (see the left panel of
Figure~\ref{Devel_parmod} for a schematic illustration). Over the last decades
this has been extended in various directions including:
\begin{itemize}
  \item \emph{Generalized linear models} (GLMs, \citealp{Nelder+Wedderburn:1972})
    encompassing an additional nonlinear link function for the conditional mean.
  \item \emph{Generalized additive models} (GAMs, \citealp{Hastie+Tibshirani:1986})
    allowing for smooth nonlinear effects in the covariates
    (Figure~\ref{Devel_parmod}, middle).
  \item \emph{Generalized additive models for location, scale, and shape}
    (GAMLSS, \citealt{Rigby+Stasinopoulos:2005}) adopting a probabilistic modeling
    approach. In GAMLSS, each parameter of a statistical distribution can depend
    on an additive predictor of the covariates comprising linear and/or
    smooth nonlinear terms (Figure~\ref{Devel_parmod}, right).
\end{itemize}
Thus, the above-mentioned models provide a broad toolbox for capturing different
aspects of the response (mean only vs.\ full distribution) and different types
of dependencies on the covariates (linear vs.\ nonlinear additive terms).

@@

While in many applications conditional mean regression models have been receiving
the most attention, there has been a paradigm shift over the last decade towards
distributional regression models. An important reason for this is that in many
fields forecasts of the mean are not the only (or not even the main) concern but
instead there is an increasing interest in probabilistic forecasts. Quantities of
interest typically include exceedence probabilities for certain thresholds of the
response or quantiles of the response distribution. As an example, consider
weather forecasting where there is less interest in the mean amount of
precipitation on the next day. Instead, the probability of rain vs.\ no rain
is typically more relevant or, in some situations, a prediction interval of
expected precipitation (say from the expected 10\% to 90\% quantile). Similar
considerations apply for other meteorological quantities and hence attention
in the weather forecasting literature has been shifting from classical linear
models \citep{Glahn+Lowry:1972} towards probabilistic models such as the
non-homogeneous Gaussian regression (NGR) of \cite{Gneiting+Raftery+Westveld:2005}.
The NGR typically describes the mean of some meteorological response variable through
the average of the corresponding quantity from an ensemble of physically-based
numerical weather predictions (NWPs). Similarly, the variance of the response
is captured through the variance of the ensemble of NWP outputs. Thus, the NGR
considers both the mean as well as the uncertainty of the NWP model outputs
to obtain probabilistic forecasts calibrated to a particular site.

%% late: introduce MOS and EMOS? more details about precipitation forecasting
%In ensemble model output statistics (EMOS, \cite{Gneiting+Raftery+Westveld:2005}) this idea is taken one step further. Usually acting on the assumption of a normally distributed response variable this model allows for both parameters, the mean and the variance, to depend linearly on covariates provided by NWPs. However, the restriction to model only linear effects has risen the interest in applying more flexible models such as offered by the GAMLSS framework.

@@

In summary, the models discussed so far provide a broad and powerful toolset for
parametric distributional fits depending on a specified set of additive linear or
smooth nonlinear terms. A rather different approach to capturing the dependence on
covariates are tree-based models.
\begin{itemize}
  \item \emph{Regression trees} (\citealt{Breiman+Friedman+Stone:1984}) recursively
    split the data into more homogeneous subgroups and can thus capture abrupt shifts
    (Figure~\ref{Devel_treeforest}, top left) and approximate nonlinear functions.
    Furthermore, trees automatically carry out a forward selection of covariates and
    their interactions. 
  \item \emph{Random forests} (\citealt{Breiman:2001}) average the predictions of
    an ensemble of trees fitted to resampled versions of the learning data. This
    stabilizes the recursive partitions from individual trees and hence better
    approximates smooth functions (Figure~\ref{Devel_treeforest}, top middle)
\end{itemize}
While classical regression trees and random forests only model the mean of the response
we propose to follow the ideas from GAMLSS modeling -- as outlined
in Figure~\ref{Devel_parmod} -- and combine tree-based methods with parametric distributional models,
yielding two novel techniques:
\begin{itemize}
  \item \emph{Distributional trees} split the data into more homogeneous groups with
    respect to a parametric distribution, thus capturing changes in any distribution
    parameter like location, scale, or shape (Figure~\ref{Devel_treeforest}, bottom left)
  \item \emph{Distributional forests} utilize an ensemble of distributional trees
    for obtaining stabilized and smoothed parametric predictions (Figure~\ref{Devel_treeforest},
    top right).
\end{itemize}
In the following, particular focus is given to distributional forests as a method for obtaining
probabilistic forecasts by leveraging the strengths of random forests: the ability to
capture both smooth and abruptly changing functions along with simultaneous selection of
variables and possibly complex interactions. Thus, these properties make the method particularly
appealing in case of many covariates with unknown effects where it would be challenging
to specify a distributional regression model like GAMLSS.

In weather forecasting, these properties are appealing in mountainous regions
and complex terrain where small-scale effects are not well-resolved in the NWP
outputs. As these are computed at a coarser scale, there may be abrupt local
effects within a single NWP grid cell. To illustrate this in practice,
precipitation forecasts are obtained with distributional forests at 95
meteorological stations in a mountainous region in the Alps, covering mainly
Tyrol, Austria, and adjacent areas. More specifically, a zero-censored Gaussian
distribution is employed to model 24-hour total precipitation so that the
zero-censored point mass describes the probability of no precipitation on a
given day (see Figure~\ref{axams_onestation}). Forecasts for July are
established based on data from the same month in the years 1985--2012 along with
82~covariates derived from a wide range of NWP ensemble outputs. As
Figure~\ref{axams_onestation} shows, the station-wise models yield a full
distributional forecast for each day -- here for the same day (July~24) at one
station (Axams) over four years (2009--2012) -- based on the previous 24~years
as training data. The corresponding observations conform reasonably well with
the predictions. In Section~\ref{ProbabilisticForecasting} we investigate the
performance of distributional forests in this forecasting task in more detail.
It is shown that they perform at least on par and sometimes clearly better than
three alternative zero-censored Gaussian models: a standard ensemble model
output statistics approach (EMOS, \citealp{Gneiting+Raftery+Westveld:2005}), a GAMLSS
with regressors prespecified based on meteorological expertise following
\cite{Stauffer+Umlauf+Messner:2017}, and a boosted GAMLSS
\citep{Hofner+Mayr+Schmid:2016} using non-homogeneous boosting
\citep{Messner+Mayr+Zeileis:2017} as a technique for variable selection from all
82~available regressors.

@@

\subsection{Motivation}
As illustrated in Figure~\ref{Devel_parmod} a wide range of tools have been developed in the field of parametric modeling (which can be applied in research areas such as probabilistic forecasting). The shift from modeling only the expected value of the response to probabilistic modeling allows for much broader inference. In many applications further features such as the ability to capture non-additive effects or interactions are desirable which demands for more flexible models. Since tree-structured models provide these required features combining them with parametric models can offer a framework that is able to deal with a wide variety of different modeling situations.

In this section the novel methodology which embeds parametric modeling into the idea of tree-structured models will be introduced step by step. Starting with a simple distributional fit in Section~\ref{distfit} will be followed by building a distributional tree in Section~\ref{disttree} and finally a distributional forest in Section~\ref{distforest}.

@@

The goal is to fit a distributional model $D(Y, \theta)$ to a response variable $Y \in \mathcal{Y}$ for a \textit{k}-dimensional vector of distribution parameters $\theta \in \Theta$ in case of a \textit{k}-parametric distribution. The distribution family has to be specified in advance such that a log-likelihood function $\ell(\theta; Y)$ is provided. Using the GAMLSS framework (\citealt{Rigby+Stasinopoulos:2005}) a wide range of distributions can be selected where each distribution parameter can be specified separately. This enables the model to deal with various features that can usually be very challenging in model fitting such as censoring, truncation or heavy tails.

In our application of precipitation forecasting the response variable, which is the power transformed total precipitation amount within 24 hours, is expected to be normally distributed with mean/location parameter $\mu$ and standard deviation/scale parameter $\sigma$ and left-censored with censoring point 0. (The power transformation is used to remove large parts of the skewness of the data as further explained in Section~\ref{data}.) Therefore, the corresponding log-likelihood function with parameter vector $\theta = (\mu, \sigma)$ is
$$
\ell(\mu, \sigma; Y) = 
\begin{cases}
    \log\left(\frac{1}{\sigma} \cdot \phi\left(\frac{Y - \mu}{\sigma}\right) \right), & \text{if } Y > 0\\
    \vspace{0.1cm}
    \log\left(\Phi\left(\frac{-\mu}{\sigma}\right)\right), & \text{if } Y = 0
\end{cases}
$$
where $\phi$ and $\Phi$ are the probability density function and the distribution function of the standard normal distribution $\mathcal{N}(0,1)$. Next to the censored normal distribution the censored logistic distribution is also often chosen for the modeling of precipitation amounts. In \cite{Gebetsberger+Messner+Mayr:2017} this choice has been made in order to be able to deal with heavy tails which often appear in precipitation data. Another frequent application of probabilistic forecasting in meteorology is the modeling of near surface air temperature. In this case a normal distribution is often selected.

@@

A common way of including these covariates is to consider them as regressors in a distributional model such that each parameter $\theta$ depends on covariates $\bold{Z} = (Z_1,\ldots,Z_m)$. This could for example be done by a linear function $g$. In that way the fitted model would be specified as $D(Y, \theta = g(\bold{Z}))$. In order to avoid the restriction to linear effects one might also use a set of smooth functions $\{f_i\}_{i = 1,\ldots,p}$ to specify a more flexible model $D(Y, \theta = f_1(\bold{Z}) + \ldots + f_p(\bold{Z}))$ applying a GAM structure. Another approach to distributional regression is proposed by \cite{Klein+Kneib+Lang:2015} where each parameter of a possibly complex distribution can be modeled by an additive term consisting of a variety of different functional effect types. These include non-linear effects, spatial effects, random coefficients and interaction surfaces which extends the flexibility compared to parametric models such as GAMLSS modeling only smooth effects.

@@

\subsection{Distributional forest}
\label{distforest}
A distributional forest consists of an ensemble of $T$ distributional trees. Each of these trees is built on a different data set which can either be a bootstrap or a subset of the original data set. Moreover, in each node a subset of the partitioning variables is chosen randomly as candidates for splitting variables. In that way the correlation among the resulting trees is reduced and therefore also the variance of the model (\citealt{Breiman:2001}).

Forest models differ mainly in the way how the trees are combined to make predictions. Following the strategy explained and proposed by \cite{Hothorn+Zeileis:2017} we interpret random forests as adaptive local likelihood estimators as distributional models are fit locally in subgroups while the term ``adaptive'' refers to the dependence on the splitting variables $\bold{Z}$ and the resulting subgroups and weighing schemes. In this case the forest itself is only used to evaluate neighborhood relationships among the observations which can then be included in the fitting process. In transformation forests by \cite{Hothorn+Zeileis:2017} so-called ``nearest neighbor weights'' are calculated as a measure of distance between observations. This approach has also already been applied in several random forest-type methods such as in ``adaptive nearest neighbors'' by \cite{Lin+Jeon:2006}, in ``quantile regression forests''  by  \cite{Meinshausen:2006} and in ``bagging survival trees'' by \cite{Hothorn+Lausen+Benner:2004}. The idea of nearest neighbor weights is to measure similarity of a pair of observations by counting the number of trees in which it is asigned to the same terminal node.
%Similar observations have a high probability of ending up in the same terminal node whereas this probability is low for quite different observations. 
%although the first random forest-type algorithm for the estimation of conditional distribution functions was published more than a decade ago (bagging survival trees, Hothorn et al. 2004). 
In particular, for a new set of covariates $\bold{z} = (z_1, \ldots, z_m) $ such a set of weights 
$\bold{w}^{\text{forest}}~=~(w^{\text{forest}}_1, \ldots, w^{\text{forest}}_n)$ is obtained by the following procedure: For each tuple $(y_i,\bold{z}_i)$ in the learning data set $\{(y_i,\bold{z}_i)\}_{i=1\ldots,n}$ a weight is calculated by counting the number of trees in which $(y_i,\bold{z}_i)$ ends up in the same terminal node as the new set of covariates $\bold{z}$:
$$
w^{\text{forest}}_i(\bold{z}) = \sum_{t=1}^T \sum_{b=1}^{B^t} I((\bold{z}_i \in \mathcal{B}^t_b) \land (\bold{z} \in \mathcal{B}^t_b))
$$
where $T$ is the number of trees of the forest, $B^t$ is the number of terminal nodes of the \textit{t}-th tree and $\mathcal{B}^t_b$ represents the \textit{b}-th terminal node of the \textit{t}-th tree.

@@

As explained in the introduction, distributional forests provide a new approach in probabilistic forecasting by embedding the ideas of flexible parametric modeling into the framework of tree-structured models. The ability to deal with a wide variety of effects and an automatic variable selection and detection of interactions are features provided by distributional forests that could be particularly valuable in probabilistic weather forecasting in order to tackle the above mentioned challenges.
To show their strengths in practice, distributional forests are now applied to probabilistic precipitation forecasts in a mountainous region based on a large number of numerical weather prediction quantities. Their results are compared to those of other zero-censored Gaussian models including a standard ensemble model output approach and two GAMLSS models. For one of the GAMLSS models the regressors have been selected and interactions have been specified in advance based on subject-matter knowledge and experience from meteorology. The other one is based on a computationally more demanding boosting approach.

@@

Observations are available on a daily basis, each representing the total precipitation sum fallen within 24~hours observed at 6~UTC at 95 stations all over Tyrol and its close surrounding. The observations are undergoing a rigorous quality check before made available to the public.  As covariates the second generation reforecast data set of the global ensemble forecast system (GEFS) is used (\citealt{Hamill+Bates+Whitaker:2013}).  This data set consists of an 11-member ensemble based on a fixed version of the numerical model and a horizontal grid-spacing of about $50~\times~50 \text{km}^2$ initialized daily at 0~UTC from December 1984 to present providing forecasts on 26 hourly temporal resolution. Each of the 11 ensemble members uses slightly different perturbed initial conditions trying to predict the situation specific uncertainty of the atmospheric state.

@@

%%% <TH> hm, now we _are_ using a transformation model without estimating the
%%% transformation. </TH>
To remove large parts of the skewness of precipitation data, a power transformation has been used frequently (Box and Cox 1964). In literature, cubic (\citealt{Stidd:1973}) or square root (\citealt{Hutchinson:1998}) transformations have often been suggested but may vary for different climatic zones or temporal aggregation periods. In \cite{Stauffer+Mayr+Messner:2017} optimization methods have been applied to find the best transformation factor which resulted in a value of $\frac{1}{1.6}$. Since the observation data set used for this optimization contains the observations considered in this article the same factor is chosen here for the power transformation of the response variable \emph{tp} (total precipitation). The resulting transformed variable is expected to be normally distributed and left-censored at censoring point 0. For simplicity the same transformation has been chosen for the variable \emph{cape} (convective available potential energy).

@@

\item An ensemble model output statistics (EMOS, \citealp{Gneiting+Raftery+Westveld:2005}) model where the location parameter and the logarithm of the scale parameter depend linearly on the ensemble mean forecast of the total daily precipitation sum and the corresponding ensemble standard deviation respectively.
The choice of the logarithmic function as link function for the scale parameter has been made based on the results of \cite{Gebetsberger+Messner+Mayr:2017} where different link functions have been compared. Moreover, testing the identity, quadratic and logarithmic function as link functions in this setting supported this choice as the logarithmic function led to the best performance of the EMOS model in this application.

\item A GAMLSS where for each distribution parameter the selection of covariates and the specification of interactions is done in advance, both based on meteorological expert knowledge following \cite{Stauffer+Umlauf+Messner:2017} (prespecified GAM). Among the available 82 regressors, 9 where selected for the location parameter and 6 for the scale parameter. Additionally, for each distribution parameter one interaction term is specified in which convective available energy is interacting with total precipitation (with the ensemble mean for the location parameter and with the ensemble standard deviation for the scale parameter). 
%In this terms an ensemble mean of 6 hour sums of convective available energy is interacting with an ensemble mean or standarad deviation of 6 hour sums of total precipitation.  
These interactions might capture events with a high potential for moderate or severe thunderstorms with large amounts of precipitation. By putting extra effort in the specification of the formulas the prespecified GAM model is provided with additional information (and is therefore expected to profit from this advantage).
\item A boosted GAMLSS \citep{Hofner+Mayr+Schmid:2016} using non-homogeneous boosting
\citep{Messner+Mayr+Zeileis:2017} for an automatic selection of covariates (boosted GAM). The idea of this (boosting) approach is to iteratively minimize a loss function where in each step only the coefficient leading to the highest improvement of the current model is updated. As the initial coefficients are all zero, only the most important covariates are included with nonzero coefficients while less relevant variables are ignored if the procedure is stopped before convergence.
\end{itemize}
